\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body} \ZZ{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}

### 2020-02-05 jfm/appendix_ams.tex
Add OU example to appendix on AMS

@@

Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
Rather, iterations were carried out until all the trajectories in the ensemble overlap.
Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
\begin{equation}
  \mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
\end{equation}
and the iterations stop.
This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "draft_p2_jfm"
%%% End:

>>>>

\section{The \ac{ams} algorithm}
\label{app_ams}

@@

Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
Rather, iterations were carried out until all the trajectories in the ensemble overlap.
Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
\begin{equation}
  \mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
\end{equation}
and the iterations stop.
This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.

\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
\label{app:AMS_on_OU}

\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
  \caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}. The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$. The blue line is the analytical solution for the return time of amplitude $a$.}
  \label{fig:comparaison_temps_de_retour}
\end{figure}
Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
\begin{equation}
  \label{eq:ou}
  \dot{x} = -x + \eta (t),
\end{equation}
where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
This basic stochastic process will allow us to highlight differences with fluid dynamics.

The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.

The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
In the latter, the typical computational cost is simply the return time $r(a)$.
One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.

Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents. Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "draft_p2_jfm"
%%% End:
